Bird beak's elaboration and exploration

### 2022-10-11 Analysis/02-mcmcmcglmmm_analysis_variance_tests.Rmd
Create 02-mcmcmcglmmm_analysis_variance_tests.Rmd

```{r, parametrise_the_model, eval = FALSE}
## Set up the parametrising chains models
param_MCMCglmm <- make.mini.chains(data         = shapespace,
                                   dimensions   = selected_dim,
                                   tree         = consensus_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "level1", "level2"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 20000,
                                       thin   = 500,
                                       burnin = 0))

## Running the three MCMCglmm models
100k_params_shapespace_8D_31levels <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(param_chains, file = "../Data/Processed/MCMCglmm/100k_params_shapespace_8D_31levels_chain1.rda")
```

@@

## Set up the mini-chains models
shapespace_allbirds_lvl_superorder_order <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 1000)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1", "level2"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_allbirds_lvl_superorder_order, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/shapespace_allbirds_lvl_superorder_order.mini.chains")
```

@@

## Checking the mini chains' ESSs

```{r, eval = TRUE}
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
pdf(file = "../Manuscript/Figures/parameters_ESS_all_birds.pdf", height = 9, width = 4)
diagnose.mini.chains(shapespace_allbirds_lvl_superorder_order[[1]])
dev.off()
```

## Passeriformes focus

## Loading the parametrisation data
load(file = "../Data/Processed/shapespace_passeriformes_two_levels.rda")

## Select the data
shapespace <- shapespace_passeriformes_two_levels[[1]]$space
selected_dim <- shapespace_passeriformes_two_levels[[1]]$dimensions
consensus_tree <- shapespace_passeriformes_two_levels[[1]]$consensus_tree
trees_list <- shapespace_passeriformes_two_levels[[1]]$trees_list
```

```{r, parametrise_the_model, eval = FALSE}
## Set up the parametrising chains models
param_chains <- make.mini.chains(data         = shapespace,
                                 dimensions   = selected_dim,
                                 tree         = consensus_tree,
                                 trait.family = "gaussian",
                                 randoms      = c("global", "level1", "level2"),
                                 residuals    = "global",
                                 priors       = 0.02,
                                 verbose      = TRUE,
                                 parameters   = list(
                                       nitt   = 50000,
                                       thin   = 500,
                                       burnin = 0))

# Running the three MCMCglmm models
save(param_chains, file = "../Data/Processed/params_passerines_shapespace_8D_28levels_50k.rda")
```

```{r, get_model_parameters, eval = FALSE}
## Load the computed models
param_chains_out <- list()
for(i in 1:3) {
    load(paste0("../Data/Cluster/shapespace_passeriformes/param_chains/params_passerines_shapespace_8D_28levels_50k_chain", i, ".rda"))
    param_chains_out[[i]] <- param_chains[[1]]
}
## Extract the parameters from the parameter chains
est_params <- extract.parameters(param_chains_out)
```

## Set up the mini-chains models
shapespace_passeriformes <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 1000)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1", "level2"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_passeriformes, file = "../Data/Cluster/shapespace_passeriformes/shapespace_passeriformes.mini.chains")
```

```{r, eval = FALSE}
## Set up the variables
models_path <- "../Data/Cluster/shapespace_passeriformes/cluster/"
## Getting all the model files
model_files <- list.files(models_path, pattern = ".rda")

## Check if the model exists already
if(!("shapespace_passeriformes.MCMCglmm" %in% list.files("../Data/Processed/"))) {
    ## Preparing the model files
    model_list <- list()
    class(model_list) <- c("mini.chains")
    ## Get the first model
    load(paste0(models_path, model_files[1], collapse = "/"))
    ## Move it to the storage
    system(paste("mv", paste0(models_path, model_files[1]), paste0(models_path, "done/")))
    print(model_files[1])
    ## Clean the bits that are not going to be used (but are huge)
    model[[1]][c("error.term", "family", "y.additional", "ginverse", "ZR")] <- NULL
    ## Update the model list
    model_list[[length(model_list)+1]] <- model[[1]]
    model <- NULL
    model_files <- model_files[-1]
} else {
    ## Updating the model
    load(file = "../Data/Processed/shapespace_passeriformes.MCMCglmm")
    model_list <- shapespace_passeriformes
}

## Combining the model files
shapespace_passeriformes <- model_list

## Saving the results
save(shapespace_passeriformes, file = "../Data/Processed/shapespace_passeriformes.MCMCglmm")
```

```{r, eval = TRUE}
load(file = "../Data/Processed/shapespace_passeriformes.MCMCglmm")
pdf(file = "../Manuscript/Figures/parameters_ESS_passeriformes.pdf", height = 9, width = 4)
diagnose.mini.chains(shapespace_passeriformes[[1]])
dev.off()
```

>>>>

Making a 2D model with just the super order random terms

```{r, parametrise_the_model, eval = FALSE}
## Set up the parametrising chains models
param_MCMCglmm <- make.mini.chains(data         = shapespace,
                                   dimensions   = selected_dim,
                                   tree         = consensus_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "level1", "level2"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 20000,
                                       thin   = 500,
                                       burnin = 0))
save(param_MCMCglmm, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k.rda")
```

## Running the parametrisation chains (for the small model)

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_2D_9levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_2D_9levels_20k, file = "../Data/Processed/MCMCglmm/params_shapespace_2D_9levels_20k_chain1.rda")
end <- Sys.time()
end - start
```

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_2D_9levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_2D_9levels_20k, file = "../Data/Processed/MCMCglmm/params_shapespace_2D_9levels_20k_chain2.rda")
end <- Sys.time()
end - start
```

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_2D_9levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_2D_9levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_2D_9levels_20k, file = "../Data/Processed/MCMCglmm/params_shapespace_2D_9levels_20k_chain3.rda")
end <- Sys.time()
end - start
```

## (for the big model)

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_8D_32levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_8D_32levels_20k, file = "../Data/Processed/MCMCglmm/params_shapespace_8D_32levels_20k_chain1.rda")
end <- Sys.time()
end - start
```

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_8D_32levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_8D_32levels_20k, file = "../Data/Processed/MCMCglmm/params_shapespace_8D_32levels_20k_chain2.rda")
end <- Sys.time()
end - start
```

```{r}
library(mcmcmcglmmm)
load(file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/params_shapespace_8D_32levels_20k.rda")
## Running the three MCMCglmm models
start <- Sys.time()
params_shapespace_8D_32levels_20k <- run.mini.chains(param_MCMCglmm, replicates = 1)
save(params_shapespace_8D_32levels_20k, file = "../Data/Processed/MCMCglmm/params_shapespace_8D_32levels_20k_chain3.rda")
end <- Sys.time()
end - start
```

@@

## Set up the mini-chains models
shapespace_allbirds_lvl_superorder_order_2D <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = 1,
                                tree         = trees_list[sample(1:1000, 1000)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_allbirds_lvl_superorder_order_2D, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/shapespace_allbirds_lvl_superorder_order_2D.mini.chains")
```

@@
